### 2018-11-08 pkg/model4you/inst/JORS/man_model4you.Rnw
comments Achim JORS
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@2382 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: 699094a147e87fe7824050639f16b5755f9a557c

\begin{document}

@@

Dependence plots with the group effect (treatment effect) on the y-axis and
the student characteristics on the x-axis are a good way of visualising the
personalised models and for getting knowledge about the interactions between
student characteristics and the treatment.
\new{
Note that the plot is related to but not the same as the classical partial
dependence plot \citep{friedman2001greedy}, as each point in the figure is one
patient in the given data set.
}
Since the percentage of successful online tests is measured on a grid, a
\new{beeswarm} plot possibly shows the relationships even better than the
scatter plot (both shown in Figure~\ref{fig:math_dp1},
\new{
Figure~\ref{fig:math_dp11} includes a loess curve). We see a nonlinear relation
between the percentage of successful online \texttt{\Sexpr{'tests'}} and the exam group
effect. Students with great performance during the semenster are estimated to
have the strongest effect.
}
<<math_dp1, fig.width=4, fig.height=3, out.width='.48\\linewidth', fig.cap="Dependence plot for percentage of tests successfully solved.", fig.subcap=c("Scatter plot with loess curve.", "Beeswarm plot.")>>=
dpdat_math <- cbind(pmods_math, MathExam)

@@

\new{
For categorical variables such as the number of previous attempts to pass the
exam (\texttt{\Sexpr{'attempt'}}) and \texttt{\Sexpr{'gender'}} we can use box plots,
\new{beeswarm} plots or a combination thereof can be used (as in
Figure~\ref{fig:math_dp21}). Figure~\ref{fig:math_dp22} shows the number of
previous attempts on the x-axis and the color indicates the gender of each
student.  This way we can see that for students writing the exam for the first
time, gender seems to play a bigger role in terms of estimated individual exam
group effect than for others.
}
<<math_dp2, fig.width=4, fig.height=3, out.width='.48\\linewidth', fig.cap="Dependence plots for the number of previous attempts and gender.", fig.subcap=c('Gender.', 'Number of previous attempts and gender.')>>=
ggplot(dpdat_math, aes(x = gender, y = group2, color = gender)) +
  geom_boxplot() +
  geom_quasirandom(alpha = 0.5) +
  theme(legend.position = "none") +
  ylab("estimated individual\nexam group effect")

@@

The software is intentionally written to make usage as simple as possible. The
most prominent use case are clinical trials where the assumption of an average
treatment effect for all patients is too strict and the efficacy of the
treatment depends on patient characteristics (e.g.~ gender, biomarkers, etc.).
For subgroup analyses (stratified treatment effects) model-based trees
(\texttt{\Sexpr{'pmtree()'}}) can be used; For personalised treatment effects
model-based forests (\texttt{\Sexpr{'pmforest()'}}) provide a way of estimating
similarity between patients and using this similarity measure to estimate
personalised models (\texttt{\Sexpr{'pmodel()'}}).  The target audience are people who
deal with heterogeneous treatment effects, such as medical researchers,
pharmaceutical companies or analysts in marketing (A-B testing).  In general
the software is useful to researchers dealing with scenarios where two
exposures are compared and responses of subjects possibly depend on other
variables.

>>>>

\begin{document}
\hypersetup{pageanchor=false}

@@

Dependence plots with the group effect (treatment effect) on the y-axis and
the student characteristics on the x-axis are a good way of visualising the
personalised models and for getting knowledge about the interactions between
student characteristics and the exam group.
\new{
Note that the plot is related to but not the same as the classical partial
dependence plot \citep{friedman2001greedy}, as each point in the figure is one
patient in the given data set.  Since the percentage of successful online tests
is measured on a grid, a beeswarm plot (a variant of jittered scatter plot)
possibly shows the relationships even better than the scatter plot (both shown
in Figure~\ref{fig:math_dp1}, Figure~\ref{fig:math_dp11} includes a loess
curve). We see a nonlinear relation between the percentage of successful online
\texttt{\Sexpr{'tests'}} and the exam group effect. Students with great
performance during the semenster are estimated to have the strongest effect.
}
<<math_dp1, fig.width=4, fig.height=3, out.width='.48\\linewidth', fig.cap="Dependence plot for percentage of tests successfully solved.", fig.subcap=c("Scatter plot with loess curve.", "Beeswarm plot.")>>=
dpdat_math <- cbind(pmods_math, MathExam)

@@

\new{
For categorical variables such as the number of previous attempts to pass the
exam (\texttt{\Sexpr{'attempt'}}) and \texttt{\Sexpr{'gender'}} we can use box plots,
\new{beeswarm} plots or a combination thereof (as in
Figure~\ref{fig:math_dp21}). Figure~\ref{fig:math_dp22} shows the number of
previous attempts on the x-axis and the color indicates the gender of each
student.  This way we can see that for students writing the exam for the first
time, gender seems to play a bigger role in terms of estimated individual exam
group effect than for others.
}
<<math_dp2, fig.width=4, fig.height=3, out.width='.48\\linewidth', fig.cap="Dependence plots for the number of previous attempts and gender.", fig.subcap=c('Gender.', 'Number of previous attempts and gender.')>>=
ggplot(dpdat_math, aes(x = gender, y = group2, color = gender)) +
  geom_boxplot() +
  geom_quasirandom(alpha = 0.5) +
  theme(legend.position = "none") +
  ylab("estimated individual\nexam group effect")

@@

The software is intentionally written to make usage as simple as possible. The
most prominent use case are clinical trials where the assumption of an average
treatment effect for all patients is too strict and the efficacy of the
treatment depends on patient characteristics (e.g.~ gender, biomarkers, etc.).
For subgroup analyses (stratified treatment effects) model-based trees
(\texttt{\Sexpr{'pmtree()'}}) can be used; For personalised treatment effects
model-based forests (\texttt{\Sexpr{'pmforest()'}}) provide a way of estimating
similarity between patients and using this similarity measure to estimate
personalised models (\texttt{\Sexpr{'pmodel()'}}).  The target audience are people who
deal with heterogeneous treatment effects, such as medical researchers,
pharmaceutical companies or analysts in marketing (A-B testing).  In general
the software is useful to researchers dealing with scenarios where two
exposures are compared and responses of subjects possibly depend on other
variables.
\new{
Currently the packages supports a limited set of model types (linear and generalised
linear models and survival models). Further models can be added. 
}
